General Relativistic Ray Tracing for Neutrinos

### 2016-10-10 paper.tex
update the abstract and create appendix

\begin{abstract}
  We present a ray tracing algorithm for computing neutrino distributions in
  general numerical spacetimes with hydrodynamical sources.
  The formulation is fully coordinate invariant and accurately describes
  radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses, and
  apply it to estimate power
  liberated by neutrino-antineutrino annihilation near the hypermassive neutron
  star formed by a merger.
\end{abstract}

@@

The neutrino distribution function, $f(p_\beta; x^\alpha)$ is an invariant
quantity counting the number of neutrinos in a given momentum volume centered on
$p_\beta$ and spatial volume centered on $x^\alpha$.
The neutrino momentum is written
\begin{equation}
  \label{eqn:def_momentum}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $u^\beta$ a general observer's velocity,
$\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
and $\ell_\beta$ the direction vector subject to the constraints
$u^\alpha \ell_\alpha = 0$ and
$\ell^\alpha \ell_\alpha=1$.
Because $\ell$ is normalized and perpendicular to the observer's velocity,
it has two degrees of freedom, which we can make explicit by defining its
spatial components with respect to spherical polar angles
\begin{equation}
  \label{eq:def_direction}
  \ell_i :=
  (\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1).
\end{equation}

Neutrino radiation obeys the relativistic Boltzmann equation
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f = \mathscr{E} - \mathscr{K} f
  + \mathscr{K}_s \oint d\Omega' \phi(\ell_\alpha,\ell_\beta') f,
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum,
and $d\Omega'$ is the element of area in momentum space, with $\ell_\alpha$
ranging over the sphere.
We have introduced
$\mathscr{E}$, the invariant emissivity,
$\mathscr{K}\equiv\mathscr{K}_a+\mathscr{K}_s$, the invariant total opacity
comprised of absorption and scattering opacities, and
$\phi(\ell_\alpha,\ell_\beta')$, the elastic scattering kernel, representing the
likelihood of scattering from the direction $\ell_\beta'$ to the direction
$\ell_\alpha$, subject to the normalization
\todo{confirm this}
\begin{equation}
  \oint d\ell' \phi(\ell_\alpha,\ell_\beta') = 1.
  \nonumber
\end{equation}

@@

A neutrino trajectory obeys the geodesic equations
\begin{equation}
\label{eqn:geodesic_x}
  \frac{d x^\alpha}{d\lambda} = p^\alpha,
\end{equation}
and
\begin{equation}
\label{eqn:geodesic_p}
  \frac{d p_\beta}{d\lambda} = -\psi_{\alpha\beta}\Gamma^\alpha_{\mu\gamma} p^\mu p^\gamma,
\end{equation}
where $\psi_{\alpha\beta}$ is the metric, and
$\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
\begin{equation}
  \label{eqn:christoffel}
  \Gamma^\beta_{\alpha\gamma} =
  \frac{1}{2} \psi^{\beta\mu}
  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}).
\end{equation}

Each trajectory is parameterized by proper distance, $s$, increasing in
the direction of $\ell$. We choose $s=0$ at $x^\alpha_e$.
and $s=s_o$ at $x^\alpha_o$.
According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
parameter by $ds=p^t d\lambda$.

@@

\section{Neutrino annihilation}
The high disk mass model is described in \cite{deat2013-leakage}.

\bibliography{references}

>>>>

\begin{abstract}
  We present a ray tracing algorithm for computing high-resolution neutrino
  distributions in general numerical spacetimes with hydrodynamical sources.
  Our formulation is covariant and treats elastic scattering by incorporating
  estimates of the neutrino fields evolved in a low-order moment simulation.
  It handles radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses, and
  apply it to simulations of neutron star mergers to
  1) compute neutrino fluxes as a function of energy and emission angle, and
  2) compute the radiation pressure tensor in the funnel of the disk,
  demonstrating a way to improve the closure relations used in the truncated
  moment formalism.
\end{abstract}

@@

The neutrino distribution function, $f(p_\beta; x^\alpha)$ is an invariant
quantity counting the number of neutrinos in a given momentum volume centered on
$p_\beta$ and spatial volume centered on $x^\alpha$.
The neutrino momentum is written
\begin{equation}
  \label{eqn:def_momentum}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $u^\beta$ a general observer's velocity,
$\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
and $\ell_\beta$ the direction vector subject to the constraints
$u^\alpha \ell_\alpha = 0$ and
$\ell^\alpha \ell_\alpha=1$.
Because $\ell$ is normalized and perpendicular to the observer's velocity,
it has two degrees of freedom, which we can make explicit by defining its
spatial components with respect to spherical polar angles
\begin{equation}
  \label{eq:def_direction}
  \ell_\alpha :=
  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
\end{equation}
with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.

Neutrino radiation obeys the relativistic Boltzmann equation
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\ell_\beta) =
  C[f](x^\alpha;\varepsilon,\ell_\beta),
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum,
and $C[f](x^\alpha;\varepsilon,\ell_\beta)$
is the source term arising from interactions with the medium,
including absorption/emission, elastic scattering, inellastic scattering,
and pair annihilation/production processes.
We symbolize $C$ this way to remind us that it depends on the positon and
momentum of the neutrinos in question, and nonlocally on $f$ at other positions
and momenta.
We compute $C$ in App.~\ref{sec:source_terms}.

We make the right hand side of Eqn.~\ref{eqn:boltzmann} explicit by writing
the source term linear in $f$:
\begin{equation}
  \label{eqn:boltzmann_linear}
  \frac{d}{d\lambda}f =
  \mathscr{E}^* - \mathscr{K}^* f,
\end{equation}
where we have introduced
$\mathscr{E}^*$, the invariant emissivity, and
$\mathscr{K}^*$, the invariant opacity, and for simplicity we have suppressed
the arguments of $f$.
These coefficients account for Fermi-blocking and involve integrals over the
neutrino distribution function at other energies and angles,
$f(\varepsilon', \ell'_\beta)$.
We derive them for the relevant microphysical processes in
App.~\ref{sec:source_terms}

@@

A neutrino trajectory obeys the geodesic equation, which may be decomposed into
the coupled first-order equations
\begin{equation}
\label{eqn:geodesic_x}
  \frac{d x^\alpha}{d\lambda} = p^\alpha,
\end{equation}
and
\begin{equation}
\label{eqn:geodesic_p}
  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\mu\beta} p^\mu p_\alpha,
\end{equation}
where $\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
\begin{equation}
  \label{eqn:christoffel}
  \Gamma^\beta_{\alpha\gamma} =
  \frac{1}{2} \psi^{\beta\mu}
  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}),
\end{equation}
and $\psi_{\alpha\beta}$ is the spacetime metric.

Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
the direction of $\ell$. We choose $\lambda=0$ at $x^\alpha_e$.
and $\lambda=\lambda_o$ at $x^\alpha_o$.
According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
parameter by $ds=p^t d\lambda$.

@@

\section{Emission Spectra}
The high disk mass model is described in \cite{deat2013-leakage}.

\section{Closure Relation}

% ******************************************************************************
\appendix

\section{Source Terms}
\label{sec:source_terms}
The source terms comprising the right hand side of the Boltzmann Equation
(Eqn.~\ref{eqn:boltzmann}), can be analyzed in terms of the microphysical
processes of absorption/emission, elastic scattering, inellastic scattering,
and pair annihilation/production.

% ******************************************************************************
\bibliography{references}
